### 2020-02-21 acros.tex
Edit chp 3 and equations
fixes and expand msr chapter
edits to msr, msfr, model chapters

@@

@@

@@

@@

@@

@@

>>>>

\newacronym{CNRS}{CNRS}{Centre National de la Recherche Scientifique, the French National Centre for Scientific Research}

@@

\newacronym{DMSR}{DMSR}{Denatured Molten Salt Reactor}

@@

\newacronym{FP}{FP}{fission product}

@@

\newacronym{HALEU}{HALEU}{high-assay low-enriched uranium}

@@

\newacronym{MCFR}{MCFR}{Molten Chloride Fast Reactor}

@@

\newacronym{MPM}{MPM}{Multi-Physics Modelling}

@@

\newacronym{RANS}{RANS}{Reynolds-averaged Navier-Stokes}

### 2020-02-21 introduction.tex
edits to msr, msfr, model chapters

>>>>

However, the liquid fuel form also brought about novel computational
challenges in simulating the transient behavior of \glspl{MSR}; the
neutronics and thermal-hydraulics are more tightly coupled due to the strong
temperature reactivity coefficient, and the additional advection coupling
terms. Furthermore, we have to account for the movement of \glspl{DNP} as
they are now generated directly within the primary coolant loop. Therefore,
the choice of coupling methods for each set of physics requires careful
consideration. Existing reactor system-level codes and modeling approaches
for conventional \glspl{LWR} analysis contain simplifying assumptions in
multiphysics coupling and other areas that render the techniques unsuitable
for simulating \gls{MSR}. Thus, making minor modifications to these codes
without changing the underlying algorithms is not the best approach for
\gls{MSR} safety analysis.

In the past two decades, researchers have developed several
new tools for simulating steady-state and transient behavior in \glspl{MSR}.
Many of the earlier efforts featured simplifications in simulating
thermal-hydraulics by solving 1-D Navier-Stokes equations or using
predetermined uniform velocity fields \cite{krepel_dyn3d-msr_2007}
\cite{kophazi_development_2009}. In more recent years, there has been
significant progress towards fully coupled, spatial codes that feature
2-D axisymmetric or full 3-D models. In 2011, Cammi et al.
\cite{cammi_multi-physics_2011} performed an ``\gls{MPM}'' analysis
of a simplified 2-D axisymmetric model of a single MSBR fuel channel using
the commercial finite-element analysis software COMSOL Multiphysics. The
physics were implemented through the two-group neutron diffusion
equations, and the \gls{RANS} standard $k-\epsilon$ turbulence model, for the
neutronics and thermal-hydraulics respectively. The
authors emphasized the need for proper full coupling of the multiphysics, and
presented both steady-state and transient results in various
scenarios such as reactivity insertions, changes in pumping rate, and the
presence of periodic perturbations. This approach featured again in a later
paper by Fiorina et al. \cite{fiorina_modelling_2014} in 2014 for a 2-D
axisymmetric model of the
\gls{MSFR}. The authors presented results from the Politecnico di Milano
COMSOL-based approach, and another approach by researchers from Delft
University of Technology, where they coupled their in-house neutronics and
thermal-hydraulics codes, DALTON and HEAT respectively. With multi-group
neutron diffusion and \gls{RANS} formulations on ultra-fine meshes, both
models showed good agreement in the steady-state neutron flux, temperature,
and \gls{DNP} distributions, and in the power responses following various
accident transient initiations. Aufiero et al. \cite{aufiero_development_2014}
concurrently developed
a full-core 3-D model of the \gls{MSFR} on OpenFOAM, albeit with one-group
neutron diffusion to reduce computational load. With the 3-D model, the
authors could simulate the asymmetric reactor response to the failure of a
single pump in the sixteen-pump \gls{MSFR} configuration. The authors also
provided some quantitative data supporting the use of implicit coupling over
explicit coupling to obtain accurate solutions of the transient cases.
Recognizing the huge computational burden required for full 3-D simulations, 
later authors came up with innovative ways to alleviate this issue such as
selective geometrical reduced order modeling for various components of a
reactor based on the importance of the physical phenomena being simulated
\cite{zanetti_geometric_2015}, or using a novel, efficient method for
neutronics calculations \cite{laureau_transient_2017}.

The preceding discussion highlights the challenges faced in MSR simulation,
and the need for a highly efficient simulation tool that incorporates
implicitly coupled multiphysics with good computational scalability over
multiple processing units. This paper presents the open source MSR simulation
tool, Moltres, as a strong contender to overcoming these challenges for
simulating the MSFR. Moltres is an application code
built in the Multiphysics Object-Oriented Simulation Environment (MOOSE)
parallel finite element framework. Similar to COMSOL and OpenFOAM, it solves
the deterministic multi-group neutron diffusion and thermal-hydraulics
\glspl{PDE} simultaneously on the same mesh. It supports up to 3-D meshes and
scales well for a large number of processors.

\section{Objectives}

In this thesis, we demonstrate the coupled multiphysics modeling capabilities
of Moltres through the reference configuration of the MSFR design. Moltres
employs fully implicit coupling between the neutronics and thermal-hydraulics
governing equations, to fully account for the tight coupling expected in MSRs
due to the movement of fuel in the salt. We implemented flow using the
Navier-Stokes equations for an accurate representation of the flow profile as
opposed to imposing a predetermined velocity field.

The objectives of this thesis are: 1) study the steady-state operating
conditions with the reference \gls{MSFR} model, 2) study the transient
behavior of the \gls{MSFR} under several unprotected accident scenarios, and
identify general characteristic behaviors in MSRs and fast spectrum reactors.
The simulation results are verified against the results by Fiorina et al.
\cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014}.

** insert thesis outline here **

### 2020-02-21 model.tex
edits to msr, msfr, model chapters

Serpent 2 is a continuous-energy Monte Carlo reactor physics code under
active development at the VTT Technical Research Centre of Finland. It was
born in 2004 out of the need for better general-purpose neutron transport
codes for group constant generation in lattice geometries, and has since grown
to support more general capabilities, for an active user base of more than 500
people worldwide. Serpent is highly parallelizable due to its support for
both MPI and OpenMP parallel programming APIs, and highly validated and
verified against experimental data and well-established codes.

@@

In the context of this project, we used Serpent with the JEFF-3.1.2 nuclear
data library* to generate group constants
for Moltres. The reactor geometry is based on the axisymmetric \gls{MSFR}
neutronics benchmark geometry published in several previous papers involving
the \gls{MSFR}. The relevant group constant data are collapsed into six
neutron energy groups, provided in table *, and calculated for temperature
values of 900 K to 1200 K at 50 K intervals.

@@

Both \gls{MOOSE} and Moltres are open source codes licensed under Lesser GNU
Public License (LPGL) terms. In the case of Moltres, this was done in support
of open and transparent research practices as well as to leverage on the ease
of access for peer review and encourage continued development of the code.
Moltres is hosted on a Github repository of the same name. Github provides
version control and continuous integration testing, which ensure existing
physics capabilities are protected from potential bugs inadvertently
introduced when developing new physics.

@@

Moltres users may call for an arbitrary number of neutron energy groups as
long as they provide Moltres with the appropriate group constant data. The
number of precursor groups is also arbitrary, though usually pre-determined by
the choice of nuclear data library in the group constant generation step.
Moltres automatically interpolates the group constant data for required
temperatures using one of the many pre-defined interpolation methods available
in \gls{MOOSE}. Once again, users have the freedom to select their
interpolation method of choice. For this work, we have selected six neutron
energy groups according to the energy boundaries in table *, eight \gls{DNP}
groups as defined by the JEFF-3.1.2 library, and the spline interpolation
method.

>>>>

Serpent 2 \cite{leppanen_serpent_2014} is a continuous-energy Monte Carlo
reactor physics code under
active development at the VTT Technical Research Centre of Finland. It was
born in 2004 out of the need for better general-purpose neutron transport
codes for group constant generation in lattice geometries, and has since grown
to support more general capabilities, for an active user base of more than 500
people worldwide. Serpent is highly parallelizable due to its support for
both MPI and OpenMP parallel programming APIs, and highly validated and
verified against experimental data and other well-established codes.

@@

In the context of this project, we used Serpent with the JEFF-3.1.2 nuclear
data library \cite{oecd/nea_jeff-3.1.2_2014} to generate group constants
for Moltres. The reactor geometry is based on the axisymmetric \gls{MSFR}
neutronics benchmark geometry published in several previous papers involving
the \gls{MSFR}. The relevant group constant data are collapsed into six
neutron energy groups, and calculated for
temperature values of 900 K to 1200 K at 50 K intervals.

@@

@@

Moltres users may call for an arbitrary number of neutron energy groups as
long as they provide Moltres with the appropriate group constant data. The
number of precursor groups is also arbitrary, though usually pre-determined by
the choice of nuclear data library in the group constant generation step.
Moltres automatically interpolates the group constant data for required
temperatures using one of the many pre-defined interpolation methods available
in \gls{MOOSE}. Once again, users have the freedom to select their
interpolation method of choice. For this work, we have selected six neutron
energy groups according to the energy boundaries in table \ref{table:bound},
eight \gls{DNP} groups as defined by the JEFF-3.1.2 library, and the spline
interpolation method.
%
\begin{table}[htb!]
	\centering
	\caption{Neutron energy group upper bounds used in Serpent.}
	\begin{tabular}{c S}
		\hline
		{Group number} & {Upper bound [MeV]}\\
		\hline
		1 & 20\\
		2 & 2.2313\\
		3 & 0.4979\\
		4 & 2.47875$\times 10^{-2}$\\
		5 & 5.5308$\times 10^{-3}$\\
		6 & 7.485$\times 10^{-4}$\\
		\hline
	\end{tabular}
	\label{table:bound}
\end{table}

### 2020-02-21 msr.tex
edits to msr, msfr, model chapters

Continuous online fuel reprocessing allows operators to maintain low excess
reactivity inventories in the core as additional fuel can be added on an ad
hoc basis. Reprocessing and gas sparging systems help reduce fissile
requirements by continuously removing neutron poisons. These factors, in
addition to the strong negative temperature
reactivity coefficient, diminishes the likelihood and severity of unprotected
criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
unlikely situation where an \gls{MSR} encounters a severe runaway reaction,
we can rely on another passive safety feature, freeze plugs. A freeze plug is
a plug filled with solidified salt at the bottom of the core
\cite{elsheikh_safety_2013}. It is actively
cooled by fans or other cooling systems to keep it just below the freezing
point of the salt. When temperatures in the core exceed a certain threshold
during a dangerous transient, the freeze plug melts and the molten salt in the
core drains into a containment tank designed to keep the salt in a
sub-critical configuration. This is especially easy to achieve for
thermal-spectrum \glspl{MSR} as the absence of moderators in the containment
tank would automatically bring the multiplication factor down below unity
\cite{elsheikh_safety_2013}. \glspl{MSR} also typically have large
margin-to-boiling under nominal operating conditions so that fuel salt boiling
does not occur. Furthermore, the reactor vessel is consequently subject to
much lower stresses as \glspl{MSR} operate at near-atmospheric pressure
levels. Thus, the probability of pipe ruptures due to high pressure is low.

@@

\subsection{Sustainability}

\section{Molten Salt Fast Reactor}

The \gls{MSFR} is a reference fast-spectrum \gls{MSR} concept developed
under the \gls{EVOL} and \gls{SAMOFAR} projects. The main reactor
specifications and schematic view are shown in Table \ref{table:msfr} and Fig.
\ref{fig:msfr} respectively. Inspired by the \gls{MSBR}, the \gls{MSFR} is
intended to run primarily on a closed thorium fuel cycle with
continuous online fuel reprocessing. Several reasons motivated the omission of
graphite moderators from the \gls{MSFR} design. Graphite is susceptible to
long-term radiation damage and replacement is likely to be necessary during
the operating lifetime of the reactor. The relatively fast neutron spectrum
improves breeding ratios due to higher fission neutron yield in $^{233}$U and
lower parasitic absorption in non-fuel materials. Lastly, graphite has a
positive temperature coefficient of reactivity; eliminating graphite from the
design ensures a greater safety margin.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.6\textwidth]{MSFR}
	\caption{Schematic view of the MSFR concept \cite{serp_molten_2014}.}
	\label{fig:msfr}
\end{figure}
%
\begin{table}[htb!]
	\caption{Main specifications of the \gls{MSFR} concept
				\cite{serp_molten_2014}.}
	\centering
	\begin{tabular}{ l r }
		\hline
		Parameter & Value \\
		\hline
		Thermal/Electric output [MW$_{\text{th}}$/MW$_{\text{e}}$] & 3000 /
		1500 
		\\
		Salt volume [m$^3$] & 18 \\
		Salt fraction in core & 0.5 \\
		Number of circulation loops & 16 \\
		Nominal flow rate [kg s$^{-1}$] & 18500  \\
		Nominal circulation time [s] & 4.0 \\
		Inlet/outlet temperature [K] & 923 / 1023 \\
		Blanket volume [m$^3$] & 7.3\\
		\hline
	\end{tabular}
	\label{table:msfr}
\end{table}

In the \gls{MSFR} design, fuel salt flows upwards through a 9 m$^3$ central
core region. At the top of the core, the flow separates into 16 smaller
external loops, each of which passes through a heat exchanger before being
pumped back into the bottom of the core. Other instrumentation are situated
along the external loop for online salt reprocessing and gas sparging. The
core is surrounded axially by nickel alloy reflectors, and radially by a
toroidal blanket tank containing fertile salt for breeding. There is a layer
of boron carbide behind the blanket tanks to protect the peripheral equipment
from excessive neutron damage. In case of severe accidents, there is an
actively cooled freeze plug at the bottom of the core that melts when
temperatures exceed a certain threshold. The fuel salt would drain into a
containment vessel designed to keep it subcritical. Reactivity control under
normal operating conditions is performed by varying pump speeds to advantage
of strong thermal feedback. Coupled with the fact that there is no excess
reactivity reserve due to online fuel reprocessing, control rods are not
included in the \gls{MSFR} design.

Although the \gls{MSFR} is primarily designed to operate on the thorium fuel
cycle, it can support a range of start-up fuel and feed compositions. This
versatility is particularly important for the first few \glspl{MSFR} to be
deployed due to the lack of $^{233}$U reserves required for the initial core
loading. In general, the fuel and blanket salts are approximately composed of
eutectic mixtures of 77.5\% LiF - 22.5\% AcF$_4$, where AcF$_4$ represents
actinide fluorides such as uranium, thorium, plutonium, and other \gls{TRU}
fluorides. For an initial composition consisting of $^{232}$Th and $^{233}$U,
the benchmark value for the amount of uranium for criticality under
normal operating conditions is 2.515 mol\%. However, most code verification
studies adjust the ratio of $^{232}$Th to $^{233}$U to achieve exact
criticality at a uniform temperature of 973 K; this ensures that subsequent
neutronics and safety analyses are not affected by the difference in
k$_{\text{eff}}$ values. We performed the same exercise in this paper.

Power output of the \gls{MSFR} is rated at 3000 MW$_{\text{th}}$ and 1500
MW$_{\text{e}}$. It has a high thermal efficiency due to the high operating
temperature. \glspl{MSR} in general are not restricted by the same pressure
constraints seen in \glspl{LWR}. The inlet and outlet temperature
specifications of the fuel salt are 923 K and 1023 K respectively. This was
motivated by the need for a minimum 50 K temperature buffer between the
operating temperatures and the melting point of the salt. The \gls{MSFR} has
heat exchangers and an intermediate coolant loop to isolate the power
conversion system from the highly radioactive fuel salt. This also serves as
a layer of containment between the radioactive material and the outside
environment. The exact composition of the intermediate coolant is still under
active study and not finalized yet.

\subsection{Model Geometry}

In this thesis, we study the 2-D axisymmetric representation of the
\gls{MSFR} as shown in Fig. \ref{fig:msfrgeom}. This is the same model
geometry used in a neutronic benchmark
study of the \gls{MSFR} and the coupled analyses by Fiorina et al., Aufiero
et al., and Zhang et al. Even though this is an outdated design that had been
later optimized for better flow distribution *, the analysis results are still
valid for code verification purposes. Future efforts will adopt the optimized
hourglass-shaped design for a more realistic assessment of the \gls{MSFR}.
%
\begin{figure}[htb!] 
	\centering
	\includegraphics[width=0.5\textwidth]{msfr-geom}
	\caption{2-D axisymmetric model of the \gls{MSFR} core used for the
	simulations in Serpent and Moltres. All dimensions are in meters.
	\cite{brovchenko_neutronic_2019}}
	\label{fig:msfrgeom}
\end{figure}

For the 2-D axisymmetric model, the sixteen individual loops are homogenized
into a single loop while keeping the nominal volumetric flow rate constant.
Keeping with the model used in the previous studies, we assumed that the
region for heat exchangers and pumps on the loop consists of 100\% fuel salt.
However, this assumption has a negligible impact on the multiplication factor
due to its location behind the strong boron carbide absorber layer;

>>>>

Continuous online fuel reprocessing allows operators to maintain low excess
reactivity inventories in the core as additional fuel can be added on an ad
hoc basis. Reprocessing and gas sparging systems help reduce fissile
requirements by continuously removing neutron poisons. These factors, in
addition to the strong negative temperature
reactivity coefficient, diminishes the likelihood and severity of unprotected
criticality accidents in \glspl{MSR} \cite{elsheikh_safety_2013}. In the
unlikely situation where an \gls{MSR} encounters a severe runaway reaction,
we can rely on another passive safety feature, freeze plugs. A freeze plug is
a plug filled with solidified salt at the bottom of the core
\cite{aji_experimental_2020}. It is actively
cooled by fans or other cooling systems to keep it just below the freezing
point of the salt. When temperatures in the core exceed a certain threshold
during a dangerous transient, the freeze plug melts and the molten salt in the
core drains into a containment tank designed to keep the salt in a
sub-critical configuration. This is especially easy to achieve for
thermal-spectrum \glspl{MSR} as the absence of moderators in the containment
tank would automatically bring the multiplication factor down below unity
\cite{elsheikh_safety_2013}. \glspl{MSR} also typically have large
margin-to-boiling under nominal operating conditions so that fuel salt boiling
does not occur. Furthermore, the reactor vessel is consequently subject to
much lower stresses as \glspl{MSR} operate at near-atmospheric pressure
levels. Thus, the probability of pipe ruptures due to high pressure is low.

@@

\subsection{Other factors}

Other factors for assessing a reactor design include sustainability,
economics, waste management, and non-proliferation.

Breeder \gls{MSR} designs easily score well in the sustainability category.
\glspl{MSR} generally have good neutron economy due to little structural
material in the core and continuous online removal of neutron poisons. Noble
metal neutron poisons also naturally plate out onto the inner walls of the
loop which are generally regions of lower neutronic importance. Both
$^{232}$Th/$^{233}$U and $^{238}$U/$^{239}$Pu fuel cycles are viable
candidates for breeding in \gls{MSR}, with the former being more suited for
thermal reactors and the latter being more suited for fast reactors. Another
benefit of this is that some \glspl{MSR} can potentially start with an initial
core loading of $^{239}$Pu with a fast spectrum before transitioning towards
$^{233}$U fuel with a $^{232}$Th feed. This measure effectively circumvents
the issue of having insufficient $^{233}$U inventory to start up the reactor.

The $^{232}$Th/$^{233}$U fuel cycle produces significantly less \gls{TRU}
waste than the other cycles due to the smaller atomic masses of $^{232}$Th and
$^{233}$U. This reduces the overall radiotoxicity and long-term decay heat
associated with long-lived plutonium and \gls{MA} isotopes. However, there are
other radionuclides such as $^{231}$Pa, $^{229}$Th, and $^{230}$U that may
pose long-term radiological concerns. Nevertheless, this feature complements
\gls{TRU} burning in fast spectrum \glspl{MSR} to reduce overall levels
of \gls{TRU} waste going into long-term storage in nuclear waste repositories.

There are some nuclear non-proliferation concerns with the thorium fuel cycle
in \glspl{MSR}.
The main concern involves the separation of the intermediate $^{233}$Pa
isotope from the fuel salt. $^{233}$Pa decays into $^{233}$U with a half-life
of approximately 27 days, and $^{233}$U is at least as potent as $^{235}$U for
nuclear weapons production. The highly radioactive $^{232}$U provides some
level of proliferation resistance but it can be sidestepped by
separating away its $^{232}$Pa precursor, which has a half-life of only 1.31
days. Strong regulatory framework and close monitoring of \glspl{MSR} would be
essential to ensure non-proliferation.

The economic prospects are uncertain at this stage of \gls{MSR} development
but we can qualitatively list the potential impacts. Capital costs are
expected to be greater due to the need for increased corrosion resistance and
the supporting reprocessing facilities. This is in spite of the reduced
pressure constraints on the reactor pressure vessel for \glspl{MSR} resulting
from the significantly lower operating pressure. There are potential cost
savings from eliminating fuel pellet and fuel assembly fabrication. Higher
fuel utilization is achieveable as neutron poisons are continuously removed
from the core via the continuous online reprocessing and gas sparging systems.
Online fuel reprocessing also minimizes reactor downtime because refuelling
can be performed at any time. The higher operating temperature results in
greater thermal efficiencies and presents new opportunities beyond
conventional electricity generation, namely industrial process heat
applications and hydrogen production.

Beyond these factors, \glspl{MSR} still require significant R\&D efforts and
engineering demonstrations for experimental validation of various components
before a full commercial model can be commissioned. Work towards creating
safety and licensing framework for \glspl{MSR} have picked up pace only in
recent years due to the growing interest from commercial \gls{MSR} developers.
